---
title: "Find well angles given MD and TVD"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment=NA, error=T, warning=F, message = F, fig.align = 'center')
```


# Challenge
Given this MD, TVD data, build a new table with angles in radians and degrees.

    0	    0
    600	    600
    1005	1000
    4075	4000
    7700	7500
    9275	9000

### Routine in Python that calculates angle and displacement

    def get_full_deviation_survey(self):
        """
        dataframe with calculations for angle and cumulative displacement
        :return: a dataframe
        """
        df = self.get_deviation_survey()
        df['tvd2'] = df.tvd - df.tvd.shift(1)
        df['md2']  = df.md  - df.md.shift(1)

        df['angle_radians'] = np.arccos(df.tvd2/df.md2)
        df['angle_degrees'] = df.angle_radians * 180 / math.pi
        df['disp'] = df.md2 * np.sin(df.angle_radians )
        df['cum_disp'] = df.disp.cumsum()
        df.fillna(0, inplace=True)
        return df
     

### Example of table after calculations

           	  MD    TVD      CumDisp     Angle
         1	 0	     0	     0	         0	
         2	 600	 600	 0	         0	
         3	 1005	 1000	 63.4429	 9.01245	
         4	 4075	 4000	 715.286	 12.2587	
         5	 7700	 7500	 1659.02	 15.0902	
         6	 9275	 9000	 2139.25	 17.7528	
   
\   
\   
\   

# <u>Solution</u>


## Read the MD, TVD table into a dataframe

```{r chunk2}
well <- "
MD	TVD
0	    0
600	    600
1005	1000
4075	4000
7700	7500
9275	9000
"

# read string text to dataframe
well_table <- read.table(header = TRUE, text = well)
well_table
```


```{r}
# a dataframe column is a vector
md2 <- well_table[, "MD"]
md2
```

## Building functions to subtract $TVD_{i+1}$ from $TVD_i$  

```{r chunk3}
# function required to move a row up or down

shift <- function(x, n, invert=FALSE, default=NA){
  stopifnot(length(x)>=n)
  if(n==0){
    return(x)
  }
  n <- ifelse(invert, n*(-1), n)
  if(n<0){
    n <- abs(n)
    forward=FALSE
  }else{
    forward=TRUE
  }
  if(forward){
    return(c(rep(default, n), x[seq_len(length(x)-n)]))
  }
  if(!forward){
    return(c(x[seq_len(length(x)-n)+n], rep(default, n)))
  }
}

# function to fill NAs with zeroes
na.zero <- function (x) {
    x[is.na(x)] <- 0
    return(x)
}
```



```{r}
# Example
shift(1:10, 5)
```

```{r}
# Example
shift(1:10, -5, default = 999)
```

## Finally, find the angle

        df['angle_radians'] = np.arccos(df.tvd2/df.md2)
        df['angle_degrees'] = df.angle_radians * 180 / math.pi
        df['disp'] = df.md2 * np.sin(df.angle_radians )

```{r}
# create table with delta MD, TVD, radian, degrees

# calculate delta MD
sh <- 1
.md2 <- well_table[, "MD"]
md2 <- shift(.md2, sh, default = 0)
well_table[, "delta.md"] <- well_table[, "MD"] - md2

# calculate delta TVD
.tvd2 <- well_table[, "TVD"]
tvd2 <- shift(.tvd2, sh, default = 0)
well_table[, "delta.tvd"] <- well_table[, "TVD"] - tvd2

# calculate the angle for each of the segments
well_table[, "radians"] <- acos(well_table[, "delta.tvd"] / well_table[, "delta.md"])
well_table[, "degrees"] <- well_table[, "radians"] * 180 / pi

# fill NAs with zeroes
well_table[, "radians"] <- na.zero(well_table[, "radians"])
well_table[, "degrees"] <- na.zero(well_table[, "degrees"])

well_table
```


