---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include = FALSE, error=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      comment = "#>",
                      collapse = TRUE,
                      error = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = 'center'
                      )
```


# Testing new deviation survey functions

## Two points
```{r}
library(rNodal)
library(dplyr)

geothermal_2p_txt <- "
    TVD   temp
    0     120
    3590  150
"

deviation_survey_2p_txt <- "
    TVD   MD
    0     0
    3590  3590
"

# convert text table of geothermal gradient to dataframe
geotherm <- as_dataframe_geothermal_data(geothermal_2p_txt)
geothermal_2p_df <- geotherm$calc_geotherm_df
geothermal_2p_df

deviation_survey_2p <- calc_deviation_survey(deviation_survey_2p_txt, 
                                              reference = "horizontal")
deviation_survey_2p
```

## Three points
```{r}
library(rNodal)
library(dplyr)

geothermal_3p_txt <- c("
  TVD    temp
  0      120
  800    129
  2670   150
")

deviation_survey_3p_txt <- "
     TVD    MD
       0     0
     800   800
    2670  2670
"

deviation_survey_3p <- calc_deviation_survey(deviation_survey_3p_txt, 
                                              reference = "horizontal")
deviation_survey_3p

# convert text table of geothermal gradient to dataframe
geotherm <- as_dataframe_geothermal_data(geothermal_3p_txt)

geothermal_3p_df <- geotherm$calc_geotherm_df
geothermal_3p_df
```

### Build iteration table (two geothermal points)

```{r table-iter_table, rows.print=25}
geothermal_txt <- geothermal_2p_txt
deviation_survey_txt <- deviation_survey_2p_txt

deviation_survey <- deviation_survey_2p
geothermal_df    <- geothermal_2p_df

iter_table <- build_iteration_table(deviation_survey, 
                                    geotherm_df = geothermal_df, 
                                    depth_points = 16)
iter_table
```

### Fluid temperature parameters using function
```{r}
# the standard well input
# Note that U (heat transfer coefficient is added at the end)

input_example_c44 <- setWellInput(
                              field.name = "HAGBR.MOD",
                              well.name = "Brown_C44", 
                              depth.wh = 0, depth.bh = 3590, 
                              diam.in = 1.995, 
                              GLR = 1000, liq.rt = 600, wcut = 0.0, 
                              thp = 500, tht = 120, bht = 150, 
                              API = 42, oil.visc = 1.0, 
                              gas.sg = 0.65, wat.sg = 1.07, if.tens = 30,
                              geotherm = geothermal_2p_txt,
                              dev_survey = deviation_survey_2p_txt
                              )

fluid_temp_params <- get_fluid_temp_parameters(input_example_c44)
fluid_temp_params
```


# Calculate fluid temperatures, bottom to top

# Simplify the Ti calculation

  (Tei - geo_grad.nz * L * sin(theta.nz)) +
    (xTi * exp(-L/A) - Tei * exp(-L/A) +
    geo_grad.nz * A * sin(theta.nz) * (1 - exp(-L/A))
  
  xTi  =  -( (Tei - geo_grad.nz * L * sin(theta.nz)) -
    Tei * exp(-L/A) +
    geo_grad.nz * A * sin(theta.nz) * (1 - exp(-L/A)) ) / exp(-L/A)
    
  xTi * exp(-L/A) + (Tei - geo_grad.nz * L * sin(theta.nz)) -
    Tei * exp(-L/A) +
    geo_grad.nz * A * sin(theta.nz) * (1 - exp(-L/A)) 
    
  xTi = xTi * K1 + K2
    

  xTi = - ( (Tei - geo_grad.nz * L * sin(theta.nz)) +
     - Tei * exp(-L/A) +
    geo_grad.nz * A * sin(theta.nz) * (1 - exp(-L/A)) ) / exp(-L/A)


```{r function-that-returns-Ti, rows.print=20}
# what if we send a smaller dataframe with only the necessary columns
# the, the table is inverted in a function that returns only the 
# calculated column
# iter-table is produced by a function

# constants
U         <- fluid_temp_params$U
diam_ft   <- fluid_temp_params$diam.ft
mass_rate <- fluid_temp_params$mass.rate
cp_avg    <- fluid_temp_params$cp.avg
k         <- U * pi * diam_ft / mass_rate / cp_avg
A         <- 1 / k          # relaxation distance by Ramey. Shoham, pg 297

fT <- function(Ti, Tei, geo_grad.nz, L, theta.nz) {
  (Tei - geo_grad.nz * L * sin(theta.nz)) +
    (Ti - Tei) * exp(-L/A) +
    geo_grad.nz * A * sin(theta.nz) * (1 - exp(-L/A))
}

calculate_Ti <- function(df) {
  df %>% 
    mutate(L = tvd[[n()]] - tvd) %>% 
    mutate(Tei = temp) %>% 
    mutate(theta = radians) %>% 
    arrange(desc(tvd)) %>% 
    mutate(theta.nz = ifelse(tvd == 0, lag(theta), theta)) %>%
    mutate(geo_grad.nz = ifelse(tvd == 0, lag(geo_grad), geo_grad)) %>% 
    { 
      x <- head(.,1)$Tei 
      sapply(1:nrow(.), function(i) { 
        x <<- fT(x, .$Tei[i], .$geo_grad.nz[i], .$L[i], .$theta.nz[i]) # global x
        .$Ti[i] <<- x  }) 
      . } %>%  # return internal dataframe
    select(-c(L, Tei, theta, geo_grad.nz, theta.nz)) %>% 
    arrange(tvd)    # go back to original order
}

iter_table_Ti <- iter_table %>% 
  mutate(Ti = NA) %>%       # add the fluid temperature variable Ti
  do(calculate_Ti(.)) %>%   # calculate Ti
  select(new_point, tvd, md, Ti, everything()) %>% 
  print()
```
